Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 300\,pc Halo

### 2021-02-26 cluster_paper/ms.tex
HR diagrams: making them awesome

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Reported components of NGC\,2516 in position and
  velocity space.} The core, identified by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, is coincident with
  the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo was identified by
  \citet{kounkel_untangling_2019} using a less restrictive membership
  assignment algorithm (see Appendix~\ref{app:clustering}).  The field
  sample is randomly drawn from a $(\alpha, \delta, \pi)$ cube
  centered on the cluster.  It is difficult to visually distinguish
  the field and the halo samples due to the low volume density of the
  halo stars.
  \label{fig:gaia6d}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			\includegraphics[width=0.85\textwidth]{f2a.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagram of NGC\,2516.}
    The photometric data are from Gaia EDR3; no reddening correction
    has been applied.  The core (black) shows a main-sequence and
    turnoff consistent with stars of a fixed age and metallicity.  The
    halo (blue) is similar, though the binary sequence is less
    defined.  At fixed color, the faintest M dwarfs in the core and
    halo are brighter than in the field star comparison sample (gray),
    consistent with these stars having not yet reached the
    main-sequence.
    \label{fig:hr}
  }
\end{figure*}

The first check on the plausibility of the candidate cluster members
being coeval was already performed by
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and more recently by
\citet{meingast_2021}.  That check is to see whether the HR diagrams
of the cluster members support the claim that they are coeval.

Figure~\ref{fig:hr} presents similar results to what these
investigators have already found.  The core members of the cluster
show a clean sequence consistent with stars with a fixed age and
metallicity, and varying mass.  The halo members show a roughly
consistent main-sequence, but with somewhat greater scatter.  One
possible explanation for this scatter is that the halo is more
contaminated by field stars.  Another possible explanation is
differential reddening.  The halo is reported to span 20$^\circ$
on-sky, and varies in position from about $b=-12^\circ$ to
$b=-20^\circ$, with the stars closest to the galactic plane also being
further from the Sun by up to 200\,pc (Figure~\ref{fig:gaia6d}).
Empirical evidence for differential reddening is discussed further in
Appendix~\ref{app:diffred}. A priori, both field star contamination
and differential reddening could well play a role.

%TODO 
Comparing these HR diagrams to the PARSEC isochrone models, we find
that X, Y, and Z.
In particular, ``the faintest M dwarfs in the core and halo are
brighter than in the field star comparison sample, consistent with
these stars having not yet reached the ZAMS.''
This is consistent with the 
main-sequence turn-off being at $Bp-Rp\approx0.05$, which implies an
age of XXX.
The resulting photometric age we calculate for the core is XXX.
For the halo, the claimed age from photometry is YYY.
Applying the same procedure to the field star comparison sample,
we get an age of ZZZ.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			\includegraphics[width=0.75\textwidth]{f2b.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagram of NGC\,2516, using Gaia EDR3 photometry.}
    {\it Bottom:} Reported members of the halo, as a function of
    galactic latitude. Can the additional scatter in the halo be
    understood through differential reddening?
    {\bf Maybe}.
    \label{fig:diffred}
  }
\end{figure*}

>>>>

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf NGC\,2516 has multiple reported components in position and
  velocity space.} The core (black) was identified by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident with
  the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo was identified by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021} using less restrictive membership
  assignment algorithms (see Appendix~\ref{app:clustering}).  The field
  sample is randomly drawn from a $(\alpha, \delta, \pi)$ cube
  centered on the cluster.  The low volume density of the
  halo stars makes it difficult to visually distinguish
  the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: The photometric data are from Gaia EDR3; no reddening correction
    has been applied.  The core (black) shows a main-sequence and
    turnoff consistent with stars of a fixed age and metallicity.  The
    halo (blue) is similar, though the binary sequence is less
    defined.  At fixed color, the faintest M dwarfs in the core and
    halo are brighter than in the field star comparison sample (gray),
    consistent with these stars having not yet reached the
    main-sequence.
    {\it Right}: Solar-metallicity PARSEC isochrones are overplotted
    at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$ with a
    reddening correction applied.
    The main-sequence turn-off location and evolved stars suggests a
    cluster age slightly older than the Pleiades.  The models diverge
    from the data for $M_\star \lesssim 0.45M_\odot$ ($\approx$M2V;
    $\mathrm{Bp}-\mathrm{Rp}\approx2.2$).
    \label{fig:hr}
  }
\end{figure*}

The first check on the plausibility of the candidate cluster members
being coeval was already performed by
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}.
That check is to see whether the HR diagrams of the cluster members
support the claim that they are coeval.

Figure~\ref{fig:hr} presents similar results to what these
investigators have already found.  The core members of the cluster
show a clean sequence consistent with stars with a fixed age and
metallicity, and varying mass.  The halo members show a roughly
consistent main-sequence, perhaps with greater scatter.  One
possible explanation for the additional scatter could be that the halo
is more contaminated by field stars.  
For instance, $\approx5$ evolved halo stars are visible in the
field-star portion of the red giant branch (RGB).
Given the expected age of the system, these stars must be field
interlopers.

A separate possible explanation for scatter in the halo's HR diagram
could be differential reddening.  The halo is reported to span
20$^\circ$ on-sky, and varies in position from about $b=-12^\circ$ to
$b=-20^\circ$, with the stars closest to the galactic plane also being
further from the Sun by up to 200\,pc (Figure~\ref{fig:gaia6d}).
Some empirical evidence for differential reddening is discussed
further in Appendix~\ref{app:diffred}. Based on the available
evidence, we expect that both field star contamination and
differential reddening could play a role.
A third possibility that we have not explored is whether the binary
fraction could differ in spatially different regions of the cluster.

In the right panel of Figure~\ref{fig:hr}, we compare the observed
photometry with PARSEC isochrones
\citep{bressan_parsec_2012,chen_improving_2014,chen_parsec_2015,marigo_new_2017}.
We used the web
interface\footnote{\url{http://stev.oapd.inaf.it/cgi-bin/cmd},
\texttt{2021-02-26}, \texttt{CMD3.4}} to interpolate these isochrones
at $\log_{10}(t/\mathrm{yr})=\{8, 8.25, 8.5\}$, and then applied the
reddening correction described in Section~\ref{subsec:tess}, in the
sense of adding the $E(Bp-Rp)$ correction to the synthetic Gaia EDR3
photometry.  The data and models generally agree for masses above
approximately 0.45\,$M_\odot$.  Below this mass, at colors redder than
Bp-Rp$\approx2.2$, the agreement is worse.  The MIST isochrones show a
comparable disagreement \citep{choi_mesa_2016}.  Explanations invoking
both starspots and incomplete molecular line lists have been proposed
\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
For purposes of age-dating the cluster, we prefer to simply focus on
the main-sequence turn-off (MSTO), since it is where the models have maximal
predictive power.

Our photometric age-dating procedure is as follows.  We selected all
stars in the cluster with absolute magnitudes $M_{\rm G}>5$.  We
excluded the six\footnote{{\bf NAMES???}} halo stars that are clear
outliers from the rest of the cluster.  Using an isochrone grid
spanning $\log_{10}(t/\mathrm{yr})=\{7.75 -- 8.75\}$ with 0.01\,dex
spacing, we then numerically evaluated the $\chi^2$ fit for each
possible age.  The resulting median, 16$^{\rm th}$, and 84$^{\rm th}$
percentiles are used as summary statstics.

% NOTE: probably asymmetric.
The resulting MSTO photometric age for the core is $XXX\pm YYY$\,Myr.
For the halo, it is $XXX\pm YYY$\,Myr.  Combining the two samples
yields $XXX\pm YYY$\,Myr.  And finally, applying the same procedure to
the field star comparison sample gives an age of $XXX\pm YYY$\,Myr.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			\includegraphics[width=0.75\textwidth]{f2c.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagram of NGC\,2516, using Gaia EDR3 photometry.}
    {\it Bottom:} Reported members of the halo, as a function of
    galactic latitude. Can the additional scatter in the halo be
    understood through differential reddening?
    {\bf Maybe}.
    \label{fig:diffred}
  }
\end{figure*}
